Handbook streamFind: Backend

### 2024-11-13 README.Rmd
Documentation improvements

# StreamFind (R package)

@@

<p align="center" width="100%">
  <img width="60%" src="man/figures/logo_StreamFind.png" alt="Logo" />
</p>

The StreamFind is developed within the project ["Flexible data analysis and workflow designer to identify chemicals in the water cycle"](https://www.bildung-forschung.digital/digitalezukunft/de/wissen/Datenkompetenzen/datenkompetenzen_wissenschaftlichen_nachwuchs/Projekte/stream_find.html), which is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de). The development is carried out by the [Institut für Umwelt & Energie, Technik & Analytik e. V. (IUTA)](https://www.iuta.de), the [Forschungszentrum Informatik (FZI)](https://www.fzi.de/) and supporting partners. The goal of the StreamFind is to develop and assemble data processing workflows for different types of analytical data (e.g., mass spectrometry and spectroscopy) and to apply the workflows in different fields (e.g., environmental and quality studies of the water cycle). StreamFind aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, to promote standardization of data structure and processing, and to facilitate retrospective data evaluation. The StreamFind platform is aimed at scientists, but also at technicians, due to its comprehensive documentation, its well categorized set of integrated modular functions and its embedded graphical user interface.  
<br>
The StreamFind development is ongoing, please [contact us](mailto:cunha@iuta.de) for questions or collaboration.

```{r, echo=FALSE, out.width = '100%', fig.align='center'}
knitr::include_graphics("man/figures/logo_group.png")
```

@@

Pre-requisites for the StreamFind are the [R](https://cran.r-project.org/) software and the [RTools](https://cran.r-project.org/bin/windows/Rtools/) (only applicable for Windows users). RTools is needed for compiling the C++ code used in the StreamFind R package. Assuming that R and RTools are installed, the StreamFind R package can be installed from the GitHub repository via the [BiocManager](https://www.bioconductor.org/install/).

```{r install-streamfind, eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
BiocManager::install("odea-project/StreamFind", dependencies = TRUE)
```

@@

The StreamFind depends on other open source software to process different analytical data. For instance, for non-target screening using mass spectrometry the StreamFind uses the [patRoon](https://github.com/rickhelmus/patRoon) R package and its own dependencies. Installation instructions for patRoon and its dependencies can be found [here](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites). Consult the documentation for dependencies of other data types.

@@

>>>>

<style>
  p {
    text-align: justify;
  }
</style>

@@

<!-- <p align="center" width="100%">
  <img width="60%" src="man/figures/logo_StreamFind.png" alt="Logo" />
</p> -->

```{r, echo=FALSE, out.width = '100%', fig.align='center'}
knitr::include_graphics("man/figures/logos.png")
```

StreamFind R package is developed within the project ["Flexible data analysis and workflow designer to identify chemicals in the water cycle"](https://www.bildung-forschung.digital/digitalezukunft/de/wissen/Datenkompetenzen/datenkompetenzen_wissenschaftlichen_nachwuchs/Projekte/stream_find.html) funded by the [German Federal Ministry of Education and Research (BMBF)](https://www.bmbf.de). The development is carried out by the [Institut für Umwelt & Energie, Technik & Analytik e. V. (IUTA)](https://www.iuta.de), the [Forschungszentrum Informatik (FZI)](https://www.fzi.de/) and supporting partners. StreamFind is intended to be a platform for assembling processing workflows for different types of data (e.g. mass spectrometry and spectroscopy) with applications in different fields (e.g. environmental and quality studies of the water cycle and quality control of pharmaceuticals). StreamFind aims to stimulate the use of advanced data analysis (e.g. non-target screening, statistical analysis, etc.) in routine studies, to promote standardisation of data structure and processing, and to facilitate retrospective data evaluation. The StreamFind platform is aimed at scientists, but also at technicians due to its comprehensive documentation, its well categorised set of integrated modular functions and its embedded graphical user interface.

The StreamFind development is ongoing, please [contact us](mailto:cunha@iuta.de) for questions or collaboration.

@@

Pre-requisites for the StreamFind are the [R](https://cran.r-project.org/) software and the [RTools](https://cran.r-project.org/bin/windows/Rtools/) (only applicable for Windows users). RTools is needed for compiling C++ code used in the StreamFind R package. StreamFind also uses python scripts for some of its functionalities, so it is recommended to have the latest [python](https://www.python.org/downloads/) installed and added to the environmental variables for Windows users. Assuming that R, RTools and python are installed, the StreamFind R package can be installed from the GitHub repository via the [BiocManager](https://www.bioconductor.org/install/).

```{r install-streamfind, eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
BiocManager::install("odea-project/StreamFind")
```

@@

The StreamFind depends on other open source software to process different analytical data. For instance, for non-target screening using mass spectrometry the StreamFind uses the [patRoon](https://github.com/rickhelmus/patRoon) R package and its own dependencies. Installation instructions for patRoon and its dependencies can be found [here](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites). When a dependency is not installed, a warning message will be displayed when necessary with instructions. Consult the documentation for dependencies of other data types.

@@

## Docker Setup

The StreamFind can also be used via the [Docker](https://www.docker.com/products/docker-desktop/) container. The Docker container is a pre-configured environment with all the necessary dependencies installed.  The Docker container can be built and started with the following commands.

Build the Docker container:  
`docker build -t my-r-app .`

Start the Docker container:  
`docker run -it -p 3838:3838 -p 8787:8787 -v $(pwd):/app my-r-app`

Once the container is up, you'll be prompted to select the service you want to run:

- *Option 1:* Starts the Shiny application, accessible at http://localhost:3838.
- *Option 2:* Starts the RStudio Server, accessible at http://localhost:8787
  + Default Username is rstudio and Password is rstudio
- *Option 3:* Starts both the Shiny App and RStudio Server
